---
title: "Physical Activity"
---

```{r, include=FALSE}
source("common_code.R")

```

Students were asked to select how many days in the last week they had done any of the following exercise types: moderate, vigorous, or strength training.[^1]

```{r, fig.width=8}
tmp1 <- cbind(data.frame(Question=rep("Moderate Exercise", 8)), as.data.frame(table(bns[["q83_1"]])))
tmp2 <- cbind(data.frame(Question=rep("Vigorous Exercise", 8)), as.data.frame(table(bns[["q83_2"]])))
tmp3 <- cbind(data.frame(Question=rep("Strength Training", 8)), as.data.frame(table(bns[["q83_3"]])))
pa.data <- rbind(tmp1, tmp2, tmp3)
colnames(pa.data) <- c('Type', 'Days', 'Freq')
rm(tmp1, tmp2, tmp3)

ggplot(pa.data) +
    geom_linerange(aes(x = Days, ymin = 0, ymax = Freq, colour = Type), 
                   position = position_dodge(width = 0.6), size=1) +
    geom_point(aes(x = Days, y = Freq, colour = Type),
               position = position_dodge(width = 0.6), size=2.5) + scale_color_manual(values=c("#336699", "#000000", "#b5d43b")) +
  labs(colour='Exercise Type') + xlab('Number of Days') + ylab('Number of Students') + theme_bw() +
  theme(legend.key.size = unit(1, 'cm'), 
        legend.key.height = unit(1, 'cm'), 
        legend.key.width = unit(1, 'cm'), 
        legend.title = element_text(size=13),
        legend.text = element_text(size=11))


```

```{r, results='asis'}
tmp <- data.frame(rbind(table(bns$q83_1),
                        table(bns$q83_2),
                        table(bns$q83_3)))

tmp <- as.data.frame(apply(tmp, 1, function(x) paste0(x, " (", percent(x/sum(x), accuracy = 0.1), ")")))
colnames(tmp) <- c("Moderate Exercise", "Strength Training", "Vigorous Training")
tmp["Days"] <- 0:7
tmp <- tmp[c(4,1,2,3)]
tmp <- t(tmp)

tmp %>%
  kable() %>% kable_styling(latex_options = c("HOLD_position", "striped", "bordered", "hover"))  %>%  
    row_spec(1, bold = T, color = "white", background = CF_green) 


```

## Moderate Exercise

How often did you do **moderate-intensity** cardio or aerobic exercise (caused a noticeable increase in heart rate, such as brisk walk) for at least 30 minutes? `r print_n_reporting("q83_1")`

-   The majority of students claim to have done moderate exercise at least once in the past week (`r sum(unname(table(bns$q83_1))[2:8])` people). The general trend is that as the number of days increases, the number of students who exercised for those given number of days decreases. This seems to be the most commonly performed type of exercise from the graph above (typically the highest number of days per week out of all exercise types).

## Strength Training

How often did you do 8-10 **strength training** exercises (such as resistance weight machines) for at least 8 repetitions each? `r print_n_reporting("q83_3")`

-    Nearly half of the students claim to have not done strength training in the past week (`r unname(sort(table(bns$q83_2), decreasing = TRUE))[1]` people). The general trend is that as the number of days increases, the number of students who exercised for those given number of days decreases. Only a small number of people claim to do strength training for 4 or more days in the past week.

## Vigorous Exercise

How often did you do **vigorous-intensity** cardio or aerobic exercise (caused large increases in breathing or heart rate, such as jogging) for at least 20 minutes? `r print_n_reporting("q83_3")`

-   The majority of students claim to have done `r names(sort(table(bns$q83_3), decreasing = TRUE))[1]` days of vigorous exercise in the past week (`r unname(sort(table(bns$q83_3), decreasing = TRUE))[1]` people). The general trend is that as the number of days increases, the number of students who exercised for those given number of days decreases. Only a small number of people claim to do vigorous exercise for 4 or more days in the past week.

---

[^1]: These questions were developed based on the [Surgeon General's physical activity guidelines for adults](https://health.gov/sites/default/files/2019-09/Physical_Activity_Guidelines_2nd_edition.pdf) who recommend: 150 minutes to 300 minutes a week of moderate-intensity exercise, 75 minutes to 150 minutes a week of vigorous-intensity aerobic exercise, or an equivalent combination of moderate- and vigorous-intensity aerobic activity. They also recommend strength-training activities on 2 or more days a week.
