---
title: "CalFresh"
---

```{r, include=FALSE}
source("common_code.R")
```

# Participation

The following questions were asked in regards to criteria laid out by the CalFresh Requirements. This includes U.S. citizenship (or lawful permanent resident), receiving Cal Grant A/B, and receiving food from a campus meal plan. Household size and monthly income is also looked at and broken down into a chart to show CalFresh eligibility. Students were also determined to be eligible or not for CalFresh based on their responses to these criteria questions.

**Pick all Basic Needs Programs and Services that you participate in (mark all that apply):**

```{r}
#| fig-height: 6
#| fig-width: 8
questions <- bns %>% select(starts_with("q46_"))

questions <-
  questions %>% mutate(across(everything(), function(x)
    fct_relevel(
      x,
      "Never heard of it",
      "Heard of it, but never used it",
      "Used it before attending college",
      "Used it in the past as a college student",
      "Currently use as a college student"
    )))

names(questions) <-
  c(
    'CalFresh (also known as Food Stamps or SNAP benefits)',
    'Campus food pantry (e.g., food, toiletries, feminine products)',
    'MediCal',
    'Emergency Housing Services'
  )
plot_likert(
  questions,
  grid.range = c(1.3, 1.3),
  wrap.legend.labels = 20,
  expand.grid = FALSE,
  values = "sum.outside",
  catcount = 5,
  geom.colors = likert.pal.5,
  show.prc.sign = TRUE,
  reverse.scale = TRUE
) + theme_bw() + theme(legend.position = "top")
```

<br>

## CalFresh Criteria

*Below are a series of Yes/No questions asked in order to help determine if a student is potentially eligible for CalFresh benefits.*

```{r}
# 5 questions reduced to two, keep as is, seperate plot for each question, ...

# If US citizen or has a green card, set to "Yes"
bns$citizenship <- ifelse((bns$q38 == "Yes" | bns$q39 == "Yes"), "Yes", "No")

# If meets income and household requirements, set to "Yes"
bns <- bns <- bns %>% mutate(m_income = as.numeric(monthly_hh_income, na.rm = TRUE),
  income_eligible = ifelse(((household_size == 1 & m_income < 3) |
                           (household_size == 2 & m_income < 4) |
                           (household_size == 3 & m_income < 5) |
                           (household_size == 4 & m_income < 6) |
                           (household_size == 5 & m_income < 7) |
                           (household_size == 6 & m_income < 8) |
                           (household_size == 7 & m_income < 9) |
                           (household_size == 8 & m_income < 10)), "Yes", "No"))

# If the student meets the 20 hour a week work requirement, set to "Yes"
bns$work_eligible <- ifelse(bns$paid_work_hours >= 20 | bns$unpaid_work_hours >= 20, "Yes", "No")

bns$EOP <- "Yes"


 qs <- c('citizenship', 'income_eligible', 'work_eligible')
 rnames <- c("US Citizen or lawful permanent resident i.e. has a ‘green card’",
             "Meets income/household requirements",
             "Meets 20 hours/week work exemption")
             
             
 binary_table(qs, 'Yes', rnames, "")
```

The data suggest the majority of students meet the citizenship and income requirements, however, less than half of the surveyed students meet the 20 hours/week work requirement.

## Household Size & Income

The following two questions were asked about household size and monthly income, which were then combined into a single plot:

-   **How many people in your household do you purchase and prepare food with, including yourself? This may or may not include your roommate(s). `r print_n_reporting("household_size")`**
-   **How much is your household's gross monthly income (before taxes and not including financial aid)? `r print_n_reporting("monthly_hh_income")`**

```{r}
library(huxtable)
tmp <- as.data.frame.matrix(table(bns$monthly_hh_income, bns$household_size))
huxtable(tmp, add_rownames = "") %>% set_background_color(2:12, 2:9, "tomato") %>%
  set_background_color(2:3, 2, "darkolivegreen3") %>% set_background_color(2:4, 3, "darkolivegreen3") %>% 
  set_background_color(2:5, 4, "darkolivegreen3") %>% set_background_color(2:6, 5, "darkolivegreen3") %>% 
  set_background_color(2:7, 6, "darkolivegreen3") %>% set_background_color(2:8, 7, "darkolivegreen3") %>% 
  set_background_color(2:9, 8, "darkolivegreen3") %>% set_background_color(2:10, 9, "darkolivegreen3") %>%
  set_background_color(11:12, 9, "yellow") %>%
  set_bold(2:3, 2) %>% set_bold(2:4, 3) %>% set_bold(2:5, 4) %>% set_bold(2:6, 5) %>% set_bold(2:7, 6) %>% 
  set_bold(2:8, 7) %>%  set_bold(2:9, 8) %>% set_bold(2:10, 9) %>% 
  set_italic(4:10, 2) %>% set_italic(5:10, 3) %>% set_italic(6:10, 4) %>% set_italic(7:10, 5) %>% 
  set_italic(8:10, 6) %>% set_italic(9:10, 7) %>%  set_italic(10, 8) %>% set_italic(11, 9) %>%
  set_all_borders(brdr(3, "solid", "white")) %>% map_align(by_cols("center")) %>%
  set_width(0.75) %>% set_font_size(12) %>% set_caption("Household Size")

# New color for row 12 column 8 (not sure if the two people qualify)
# Combine rows 2 and 3?
```

*In the above graph, cells highlighted in green & bold indicate a person is potentially eligible for CalFresh based on their income, cells highlighted in red & italics indicate a person is not eligible, and cells highlighted in yellow indicate a person may be eligible. This information is pulled from the CalFresh requirements section A2 (see below for link).*

## CalFresh Eligible

**This variable was created under the following criteria `r print_n_reporting("calfresh_eligible")`**

*Disclaimer: only county eligibility workers have the authority to determine official eligibility for CalFresh. These students are just most likely to be eligible based on the below listed criteria.*

-   The person is between the ages 18 to 49.\
-   The person is at least a part-time student (6 units per semester for the three campuses in this survey).\
-   The person is either a U.S. citizen or lawful permanent resident.\
-   Does not get more than half of their meals from meals plans (less than 11 per week or none).\
-   Identifies as an EOP student.\
-   Falls under the household size/income from the [Department of Social Services](https://www.cdss.ca.gov/inforesources/cdss-programs/calfresh/eligibility-and-issuance-requirements#income) requirements.

<div class = "row">

::: col-md-6
```{r}
c1 <- data.frame(bns$calfresh_eligible) %>% na.omit()
c1 <- c1 %>% mutate(ifelse(bns.calfresh_eligible==0, "No", "Yes"))
plot_frq(c1$`ifelse(bns.calfresh_eligible == 0, "No", "Yes")`, geom.colors = plot_color,  ylim=c(0,1600)) + 
  xlab('Likely Eligible for CalFresh') + ylab("Frequency")
```
:::

We can see that the majority of students are most likely eligible for CalFresh (`r unname(table(bns$calfresh_eligible)['1'])` students), while `r unname(table(bns$calfresh_eligible)['0'])` students are most likely to not be eligible for these benefits.

<div class = "col-md-5">

# Impact

The following questions were asked in regards to the impact of Basic Needs programs and services on students lives. Students were asked to evaluate how they feel these programs have benefited them both academically and non-academically, as well as barriers/reasons in regards to why they are not using certain programs. This includes programs such as CalFresh, Food Pantries, Medi-Cal (also known as Medicaid in the United States), and Emergency Housing Services.

<br>

## CalFresh EBT

*Note: for the graph on the left (EBT application) we have `r print_n_reporting("q47")` For the graph on the right (people fed through EBT) we have `r print_n_reporting("fed_through_EBT")`*

```{r}
#| fig-cap-location: top
#| fig-subcap: 
#|   - "Did you apply for CalFresh EBT as an individual or with other people in your household?"
#|   - "How many people in the household are regularly fed through your CalFresh EBT card?"
#| layout-ncol: 2
#| fig-width: 5
p1 <- plot_frq(bns$q47, geom.colors = plot_color, wrap.title = 40, ylim=c(0,500), wrap.labels = 15) +
  xlab('')
p2 <- plot_frq(bns$fed_through_EBT, geom.colors = plot_color, wrap.title = 40) + xlab('')

p1
p2
rm(p1,p2)
```

Looking at the above graphs, the majority of students apply for CalFresh EBT as an individual (`r unname(table(bns$q47)['As an individual'])` students) while only `r unname(table(bns$q47)['With other people in my household'])` students applied with other people in their household. Similarly, the majority of students claimed only `r names(sort(table(bns$fed_through_EBT), decreasing = TRUE))[1]` person is fed through the EBT card (`r unname(sort(table(bns$fed_through_EBT), decreasing = TRUE))[1]` people). `r unname(sort(table(bns$fed_through_EBT), decreasing = TRUE))[2]` students claim to feed `r names(sort(table(bns$fed_through_EBT), decreasing = TRUE))[2]` people through the EBT card, and only a small amount of students claim to feed 5 or more people through the EBT card.

<br>

### Monthly CalFresh Money

**How much money do you receive from CalFresh each month? `r print_n_reporting("q49_1")`**

```{r}
#| fig-width: 7

gghistogram(bns, x = "q49_1", add = "mean", fill = plot_color, alpha = 1) + xlab("Monthly EBT Received ($)") + ylab("Frequency")


#is $1500 a feasible response?
```

The average amount of money received per household using CalFresh (denoted by the dotted line) is \$`r round(mean(bns$q49_1, na.rm=TRUE),2)`, while the median is \$`r round(median(bns$q49_1, na.rm=TRUE),2)`. The values range from \$`r min(bns$q49_1, na.rm=TRUE)` to \$`r max(bns$q49_1, na.rm=TRUE)`. The majority of households who do receive these benefits seems to be getting around \$250 or less per month.

### Household Size Breakdown

Below is a breakdown of the average monthly allowance received from CalFresh based on household size.

```{r}
#| fig-width: 6.5
tmp <- bns %>% group_by(fed_through_EBT) %>% 
  summarise(avg = round(mean(q49_1, na.rm=TRUE),2)) %>% 
  na.omit() %>% mutate(HH.avg = round(avg/fed_through_EBT, 2),
                       fed_through_EBT = as.character(fed_through_EBT),
                       avg = paste0("$", format(avg, nsmall=2)))

tmp[9,] <- list("", "Total Average", round(mean(tmp$HH.avg),2))
tmp <- tmp %>% mutate(HH.avg = paste0("$", format(HH.avg, nsmall=2)))
colnames(tmp) <- c("Household Size", "Household Average", "Per Person Average")

tmp %>% kable() %>% kable_styling(bootstrap_options = "striped") %>%
  row_spec(9, bold=TRUE)
```

The general pattern seems to be that as household size increases so does the monthly average assistance from CalFresh. The household average was divided by the household size to find the "per person average" for CalFresh assistance. From taking the average of these values, we can see that a person on CalFresh receives an average monthly assistance of: `r tmp[9,3]`.

### Is This Enough?

*Note: for the first graph on if the allowance is enough we have `r print_n_reporting("q50")` And for the second graph on when people run out we have `r print_n_reporting("q51")`*

```{r}
#| fig-cap-location: top
#| fig-subcap: 
#|   - "I feel that the amount that is added to my CalFresh EBT card each month is enough money to provide healthy food for the month."
#|   - "When do you typically use up the balance on your CalFresh EBT card?"
#| layout-ncol: 2
  
p1 <- plot_frq(bns$q50, geom.colors = plot_color, ylim=c(0,200)) + xlab('')
p2 <- plot_frq(bns$q51, geom.colors = plot_color, ylim=c(0, 210), wrap.labels = 12) + xlab('')

p1
p2
rm(p1,p2)
```

Of the students who do receive money from CalFresh, `r unname(table(bns$q50)['Agree']) + unname(table(bns$q50)['Strongly agree'])` of them agree that it is enough money to provide healthy food for a month and `r unname(table(bns$q50)['Disagree']) + unname(table(bns$q50)['Strongly disagree'])` of them feel that it is not enough. Students typically run out of CalFresh money in the first week (`r unname(table(bns$q51)['First week after receiving benefits'])` students), the second week (`r unname(table(bns$q51)['Second week after receiving benefits'])` students) or the third week (`r unname(table(bns$q51)['Third week after receiving benefits'])` students). A smaller minority run out within the fourth week or do not use up all their monthly balance.

<br>

## CalFresh Benefits

**While you received CalFresh benefits:**

```{r}
#| fig-height: 8
#| fig-width: 8.5
questions <- bns %>% select(starts_with("q52_"))
names(questions) <- c('I worked fewer hours each week in a PAID position.',
                      'I had more time to do an UNPAID internship or volunteer work.',
                      'I had more money to pay for bills (rent, utility bills, clothes, etc.).',
                      'I had more money for entertainment and social activities.',
                      'I shopped more frequently at the farmers market.',
                      'I had learned more about healthier foods.',
                      'I purchased healthier foods.',
                      'I cooked more frequently at home.',
                      'I consumed a healthier diet.',
                      'I consumed less fast food.',
                      'I consumed fewer processed foods.',
                      'I consumed fewer sugar-sweetened beverages like soda and energy drinks.',
                      'I consumed less alcohol or fewer alcoholic beverages.',
                      'I concentrated more in class(es).',
                      'My grades had improved.',
                      'My grades had decreased.',
                      'I kept up with course materials.',
                      'I had more time to go to my professors office hours and tutoring services when needed.',
                      'I had more time to attend all classes as they are scheduled.',
                      'I had more time to go on field trips or internships.',
                      'My sleep quality had improved.',
                      'My stress/anxiety level had decreased.')
plot_likert(questions[1:11],
            grid.range = c(1.2, 1.2),
            wrap.legend.labels = 20,
            expand.grid = FALSE,
            values = "sum.outside",
            catcount = 5,
            geom.colors = likert.pal.5,
            show.prc.sign = TRUE, reverse.scale = TRUE) + theme_bw() + theme(legend.position = "top")

plot_likert(questions[12:22],
            grid.range = c(1.2, 1.2),
            wrap.legend.labels = 20,
            expand.grid = FALSE,
            values = "sum.outside",
            catcount = 5,
            geom.colors = likert.pal.5,
            show.prc.sign = TRUE, reverse.scale = TRUE) + theme_bw() + theme(legend.position = "top")
```

## CalFresh Barriers

**What are the reasons you do not have CalFresh or have never used CalFresh? (Select all that apply).**

```{r}
question <- "calfresh_barrier"

qlabel <- c('No time to complete the application/ application is too long.', 
'The application is too difficult.', 
'The CalFresh outreach team on campus was not helpful or supportive during my application process.',
'The staff in my local county social services office were not helpful or supportive during my application process.', "It's too difficult to get my verification documents.",
'No time to call or go to the local social services office for my CalFresh interview.',
'I feel other students need CalFresh more than I do.', 
'No transportation to CalFresh office or social services office.',
'I do not think I am eligible.', 
"I have heard that it's not worth my time because I won't be eligible or I won't receive much money.", 
'My family or household does not want to apply for CalFresh.', 
'I am embarrassed or ashamed to apply for CalFresh.', 
"I don't believe in social services like CalFresh, WIC, or Welfare.",
"I don't need CalFresh.", 'Other (Please specify).', 
'I feel uncomfortable receiving help with basic needs, like food.', 
'I would think less of myself if I applied for CalFresh.',
"I don't wan't my friends to find out I'm receiving food assistance.", 
'My friends and peers would think less of me if I apply for CalFresh.')




cf_barrier_vars <- prep_binary_vars(question, qlabel) ## Issues here



# Plot table of Percentages
cf_barrier_vars %>%
select(Measure = xlab, "Yes (%)" = pct_lab) %>%
kable() %>% kable_styling(bootstrap_options = "striped") %>%  column_spec(2, width='3.5cm')
```

## Food Pantry Use

**How often do you normally use the food pantry or if you are not currently using it, how often did you use it? `r print_n_reporting("q54")`**

```{r}
#| fig-height: 3.5
plot_frq(bns$q54, geom.colors = plot_color, wrap.labels = 15) + xlab('Food Pantry Use Frequency') + ylab("Frequency")
```

When students use the food pantry, the majority do not use it every week (`r unname(sort(table(bns$q54), decreasing = TRUE)['Not every week'])` students). `r unname(sort(table(bns$q54), decreasing = TRUE)['About once a week'])` students claim to use it about once a week, `r unname(sort(table(bns$q54), decreasing = TRUE)['More than once a week'])` students use it more than once a week, and `r unname(sort(table(bns$q54), decreasing = TRUE)['Other (Please specify):'])` students claim other uses.

# Other Basic Needs Services

## Food Pantry - Impact
**For the following questions, please choose how much you agree or disagree with the impact campus food pantries has or had on your life. While I utilized campus food pantry/pantries:**

```{r}
#| fig-height: 8
#| fig-width: 8.5
questions <- bns %>% select(starts_with("q55_"))
names(questions) <- c('I worked fewer hours each week in a PAID position.',
                      'I had more time to do an UNPAID internship or volunteer work.',
                      'I had more money to pay for bills (rent, utility bills, clothes, etc.).',
                      'I had more money for entertainment and social activities.',
                      'I shopped more frequently at the farmers market.',
                      'I had learned more about healthier foods.',
                      'I purchased healthier foods.',
                      'I cooked more frequently at home.',
                      'I consumed a healthier diet.',
                      'I consumed less fast food.',
                      'I consumed fewer processed foods.',
                      'I consumed fewer sugar-sweetened beverages like soda and energy drinks.',
                      'I consumed less alcohol or fewer alcoholic beverages.',
                      'I concentrated more in class(es).',
                      'My grades had improved.',
                      'My grades had decreased.',
                      'I kept up with course materials.',
                      'I had more time to go to my professors office hours and tutoring services when needed.',
                      'I had more time to attend all classes as they are scheduled.',
                      'I had more time to go on field trips or internships.',
                      'My sleep quality had improved.',
                      'My stress/anxiety level had decreased.')
plot_likert(questions[1:11],
            grid.range = c(1.2, 1.2),
            wrap.legend.labels = 20,
            expand.grid = FALSE,
            values = "sum.outside",
            catcount = 5,
            geom.colors = likert.pal.5,
            show.prc.sign = TRUE, reverse.scale = TRUE) + theme_bw() + theme(legend.position = "top")

plot_likert(questions[12:22],
            grid.range = c(1.2, 1.2),
            wrap.legend.labels = 20,
            expand.grid = FALSE,
            values = "sum.outside",
            catcount = 5,
            geom.colors = likert.pal.5,
            show.prc.sign = TRUE, reverse.scale = TRUE) + theme_bw() + theme(legend.position = "top")
```

## Food Pantry - Barrier

**What are the reasons for you to not go to campus food pantries? (Select all that apply).**

```{r}
 question <- "no_pantry"
 qlabel <- c("My family doesn't want me to go to campus food pantries or local food banks.",
             'I am embarrassed or scared to go to the campus food pantries or local food banks.',
             'I do not believe in using campus food pantries or local food banks.',
             "Inconvenient to access due to lack of transportation or out-of-the way location.",
             'Inconvenient times of operation/limited pantry hours.',
             'I feel other students need the food more than I do.',
             'I am not aware of any campus food pantries or local food banks.',
             'There are no food pantries close to campus or my residence.',
             'Other', 
             "I don't need to use campus food pantries.", 
             'I feel uncomfortable about receiving help with basic needs, like food.',
             'I would think less of myself If I went to campus food pantries or local food banks.',
             "I don't want my friends to find out I go to campus food pantries or local food banks.",
             'My friends and peers would think less of me if I went to campus table(food pantries or local food banks.')
 
 pantry_vars <- prep_binary_vars(question, qlabel)

# Plot table of Percentages
 pantry_vars %>% 
   select(Measure = xlab, "Yes (%)" = pct_lab) %>% 
   kable() %>% kable_styling(bootstrap_options = "striped")
```


## Medi-Cal - Impact

**Knowing that you have Medi-Cal:**

```{r}
#| fig-width: 8.5
questions <- bns %>% select(starts_with("q59_"))
names(questions) <- c("My mental health has improved.",
                      "My physical health has improved.",
                      "My sleep quality had improved.",
                      "My stress/anxiety level had decreased.")
                      

plot_likert(questions,
            grid.range = c(1.2, 1.2),
            wrap.legend.labels = 20,
            expand.grid = FALSE,
            values = "sum.outside",
            catcount = 5,
            geom.colors = likert.pal.5,
            show.prc.sign = TRUE, reverse.scale = TRUE) + theme_bw() + theme(legend.position = "top")
```

## Medi-Cal - Barriers

**What are the reasons you do not have Medi-Cal or have never used Medi-Cal? (Select all that apply).**

```{r}
 question <- "no_medical"
 qlabel <- c("I don’t need MediCal.",
             "I do not know what MediCal is or I think I already have health insurance.",
             "I only need medical services at the student health center associated with my campus.",
             "No time to complete the MediCal application/MediCal application is too long.",
             "The MediCal application is too difficult to complete.",
             "I do not think I am eligible for MediCal.",
             "I have heard from other students that it is not worth my time because I wont be eligible.",
             "My family or household does not want me to apply for MediCal.",
             "I am embarrassed or ashamed to apply for MediCal.",
             "I don’t believe in social services like MediCal, CalFresh, WIC, or Welfare/Cash Aid.",
             "Other",
             "I feel uncomfortable about receiving help with basic needs, like food.",
             "I would think less of myself If I went to campus food pantries or local food banks.",
             "I don't want my friends to find out I go to campus food pantries or local food banks.",
             "My friends and peers would think less of me if I went to campus table(food pantries or local food banks.")
 
 medical_vars <- prep_binary_vars(question, qlabel)
 
 # Plot table of Percentages
 medical_vars %>% 
   select(Measure = xlab, "Yes (%)" = pct_lab) %>% 
   kable() %>% kable_styling(bootstrap_options = "striped")
```

## Emergency Housing Services - Impact

**While you received Emergency Housing Services:**

```{r}
#| fig-height: 8
#| fig-width: 8.5
questions <- bns %>% select(starts_with("q62_"))
names(questions) <- c('I concentrated more in class(es).',
                      'My grades had improved.',
                      'My grades had decreased.',
                      'I kept up course materials.',
                      'I had more time to go to my professors’ office hours and tutoring services when needed.',
                      'I had more time to attend all classes as they are scheduled.',
                      'I had more time to go on field trips or internships.',
                      'My sleep quality had improved.',
                      'My stress/anxiety level had decreased.',
                      'I’m less concerned about housing issues.',
                      'I felt more confident in my ability to maintain healthy relationships.')

plot_likert(questions,
            grid.range = c(1.2, 1.2),
            wrap.legend.labels = 20,
            expand.grid = FALSE,
            values = "sum.outside",
            catcount = 5,
            geom.colors = likert.pal.5,
            show.prc.sign = TRUE, reverse.scale = TRUE) + theme_bw() + theme(legend.position = "top")
```

## Emergency Housing Services - Barriers

**What are the reasons you do not have Emergency Housing Services or have never used Emergency Housing Services?**

```{r}
 question <- "no_ehs"
 qlabel <- c("I don’t need Emergency Housing Services",
             "I was not aware that my campus had Emergency Housing Services like grants and housing placements in an emergency.",
             "No time to complete the Emergency Housing application or contact the Emergency Housing Staff.",
             "The Emergency Housing Services application is too difficult to complete.",
             "The emergency housing grant/money is not enough to support what I need, so not worth my time to apply for it.",
             "I am embarrassed or ashamed to apply for Emergency Housing Services.",
             "I do not believe in social services like Emergency Housing Services, MediCal, CalFresh, or Welfare/Cash Aid.",
             "Other", 
             "I feel uncomfortable about receiving help with basic needs, like housing.",
             "I would think less of myself if I applied for Emergency Housing Services.",
             "I don't want my friends to find out that I applied for emergency housing services.",
             "My friends and peers would think less of me if I applied for emergency housing services.")
             
 
 ehs_vars <- prep_binary_vars(question, qlabel)
 
 # Plot table of Percentages
 ehs_vars %>% 
   select(Measure = xlab, "Yes (%)" = pct_lab) %>% 
   kable() %>% kable_styling(bootstrap_options = "striped") %>% column_spec(2, width='3.5cm')
```
