---
title: "Demographics"
---

```{r, include=FALSE}
source("common_code.R")
```

# Personal Demographics

The following set of questions were asked in regards to the students' personal demographics. This includes the students' age, ethnicities they identify with, and the number of dependents they are responsible for taking care of. They were also asked about parents' education, which will help us identify first generation college students.


## Age

*The following was determined by taking the difference between birth date and the time at which the survey was completed* **`r print_n_reporting('age')`**

```{r}
gghistogram(bns, x = "age", add = "mean", alpha = 1, 
            fill = plot_color) + 
  xlab("Age") + ylab("Number of Students")
```

The age of students in the survey range from `r min(bns$age, na.rm = TRUE)` to `r max(bns$age, na.rm = TRUE)` at the time of completion, with the majority of students being between 20-30 years old. The average age of a participant in this survey is `r round(mean(bns$age, na.rm = TRUE),1)` years old. <br />

<br>

## Gender Identity

**What is your gender identity? `r print_n_reporting('q7')`**

<div class = "row">

::: col-md-7
```{r}
#| fig-height: 3.5
#| fig-width: 7
rec <- bns %>% mutate(q7=recode(q7, "Gender non-binary"="Non-binary", "Gender Queer or Gender-Nonconforming" = "Queer or Nonconforming", "Gender Unlisted:" = "Unlisted"))


plot_frq(rec$q7, sort.frq="desc", geom.colors = plot_color) + xlab('Gender') + ylab("Number of Students")
#Should we cut "Gender" from the 3 categories on the right? Exp. "Non-binary" instead of "Gender non-binary"
```
:::

Of the students in the survey, `r unname(table(bns$q7)['Woman'])` of them identify as a woman, `r unname(table(bns$q7)['Man'])` identify as a man, and `r unname(table(bns$q7)['Gender non-binary'])` identify as gender non-binary. There are also students who preferred not to answer.

<br>

## Ethnicity

**In the US Census, Hispanic Origins are not races. Do you identify as Hispanic or Latino or Spanish Origin? `r print_n_reporting('q8')`**

::: row
::: col-md-6
```{r}
#| fig-height: 3
#| fig-width: 4
tmp <- data.frame(q8=bns$q8) %>% mutate(hispanic=as.factor(ifelse(is.na(q8), NA, ifelse(grepl("Yes", q8), "Hispanic", "Non-Hispanic")))) %>% select(-q8)

plot_frq(tmp$hispanic, geom.colors = plot_color, ylim=c(0, 1800)) + xlab("") + ylab("Number of Students")
```
:::

::: col-md-5
We define Hispanic Origin to include: Mexican, Mexican American, Chicano, Puerto Rican, or another Hispanic/Latino/Spanish origin. The data indicates that the majority of students identify as having Hispanic Origins (`r unname(table(tmp$hispanic)['Hispanic'])` students), while around a quarter of them do not identify as having Hispanic Origins (`r unname(table(tmp$hispanic)['Non-Hispanic'])` students).
:::
:::


```{r}
tmp <- data.frame(q8=bns$q8, school=bns$school) %>% 
  mutate(hispanic=as.factor(ifelse(is.na(q8), NA, ifelse(grepl("Yes", q8), "Yes", "No"))),
         school=ifelse(is.na(hispanic), NA, school)) %>% select(-q8)
```

<br>

## Race

```{r}

tmp <-
  bns %>% select(starts_with("eth_")) %>% mutate(
    white = ifelse(is.na(eth_white), NA, ifelse(eth_white == 1, 1, 0)),
    black = ifelse(is.na(eth_african_amer), NA, ifelse(eth_african_amer == 1, 1, 0)),
    amer_indian = ifelse(is.na(eth_amer_indian), NA, ifelse(eth_amer_indian == 1, 1, 0)),
    asian = ifelse((
      is.na(eth_chinese) | is.na(eth_filipino) | is.na(eth_asian_indian) |
        is.na(eth_other_asian) |
        is.na(eth_vietnamese) | is.na(eth_korean) |
        is.na(eth_japanese)
    ),
    NA,
    ifelse((
      eth_chinese == 1 | eth_filipino == 1 | eth_asian_indian  == 1 |
        eth_other_asian == 1 |
        eth_vietnamese == 1 | eth_korean == 1 |
        eth_japanese == 1),1,0)
    ),
    hawaiian = ifelse((
      is.na(eth_pacific_islander) |
        is.na(eth_chamorro) | is.na(eth_hawaiian) |
        is.na(eth_samoan)
    ), NA,
    ifelse((
      eth_pacific_islander == 1 | eth_chamorro == 1 | eth_hawaiian == 1 |
        eth_samoan == 1
    ),
    1,
    0
    ))
  ) %>% select(15, 17:21)


```

**What race do you identify as? (Select all that apply) `r print_n_reporting_tmp('white')`** *Note: that the survey included 14 race options that were collapsed down to 5 broad groups defined by [census.gov](https://www.census.gov/quickfacts/fact/note/US/RHI625219).*

```{r}
cnames <- c("Prefer not to answer race", "White", "Black or African American", "American Indian and Alaska Native", "Asian", "Native Hawaiian and Other Pacific Islander")
tmp2 <- tmp %>% na.omit()
colnames(tmp2) <- cnames
tmp2 <- as.data.frame(t(tmp2)) %>% mutate(sums=rowSums(.),
                                          label=paste0(sums, " (", percent(sums/ncol(.), accuracy=0.1), ")")) %>% arrange(desc(sums)) %>% select(label)
colnames(tmp2) <- "Yes (%)"
tmp2 %>% kable() %>% kable_styling(bootstrap_options = "striped")
```

<br>

## Dependents

**If you have dependents, then how many of them receive more than 50% of their financial support from you? `r print_n_reporting('q11')`**

```{r}
#| fig-height: 4
#| fig-width: 6
plot_frq(bns$q11, geom.colors = plot_color,) + xlab('Number of Dependents') + ylab("Number of Students")
```

The majority of students who report having at least one dependent also report that they provide more than 50% of financial support to at least one dependent (`r nrow(bns[bns$q11 > 0, 'q11'] %>% na.omit())` students).

<br>

## Parents Education

**What is the highest level of education completed by your parents/guardians (for example, if your mother and father both raised you, and your mother has the highest level of education, then indicate her highest level of education): `r print_n_reporting('q13')`**

```{r}
#| fig-height: 5
#| fig-width: 7.5
tmp2 <- as.data.frame(table(bns$q13)) %>%
  mutate(Percent=lapply(Freq, function(x) paste0('(', percent(x/sum(!is.na(bns[['q13']])), accuracy = 0.1), ')')),
         FirstGen=as.character(c(1,1,1,1,1,0,0,0,1)),
         Var1=recode(Var1, "Other/Unknown" = "Other/ Unknown")) %>% arrange((desc(Freq)))

ggplot(tmp2, aes(x=factor(Var1, levels=Var1), y=Freq, fill=FirstGen, label=Percent)) + geom_col(colour = "black") +
  scale_fill_manual(values=c("#336699", "#ecf0b7")) + geom_text(aes(y=Freq+50)) + ylim(0,900) +
  scale_x_discrete(labels=label_wrap(10)) +   geom_text(aes(y=Freq + 100, label=Freq)) +
  xlab('') + ylab("Number of Students") + theme(legend.position = "none") + theme(axis.text.x = element_text(size = 10))
```

*Note: Categories in yellow were used to determine "First Generation" status.* 

The highest level of education completed by most students' parents was "`r names(sort(table(bns$q13), decreasing=TRUE)[1])`" (`r unname(sort(table(bns$q13), decreasing=TRUE)[1])`), while `r unname(sort(table(bns$q13), decreasing=TRUE)[2])` students reported their parents highest level is "`r names(sort(table(bns$q13), decreasing=TRUE)[2])`", and `r unname(sort(table(bns$q13), decreasing=TRUE)[3])` report their parents highest level is "`r names(sort(table(bns$q13), decreasing=TRUE)[3])`". Only a minority of students (`r sum(table(bns$q13)[c("Bachelor's degree", "Master's degree", "Doctorate degree (JD, MD, PhD, EdD., etc.)")])`) indicated that their parents have a bachelor's degree or higher.

<br>

## First Generation

*Note: Students with parents education identified as "Other/Unknown" were not marked as first gen or not since they could not be determined.* **`r print_n_reporting('firstgen')`**

::: row
::: col-md-6
```{r}
#| fig-height: 3
#| fig-width: 4
tmp3 <- data.frame(firstgen=bns$firstgen) %>% 
  mutate(firstgen=factor(ifelse(is.na(firstgen), NA, ifelse(firstgen == 0, "Not First Gen", "First Gen")), levels=c("Not First Gen", "First Gen")))

plot_frq(tmp3$firstgen, geom.colors = plot_color, ylim=c(0, 2300)) + xlab('') + ylab("Number of Students")
```
:::

::: col-md-5
`r unname(sort(table(bns$firstgen), decreasing=TRUE)[1])` students are considered as First Generation College students and `r unname(sort(table(bns$firstgen), decreasing=TRUE)[2])` students are not considered First Generation College students.
:::
:::

As defined by the [University of California](https://www.universityofcalifornia.edu/sites/default/files/First-Generation%20Student%20Success%20at%20UC%208-2017.pdf) first generation students are "working to be part of the first generation in their families to earn a four-year university degree". This means based upon parents education, students in the data set are considered first generation if their parents do not have a Bachelor's or Master's degree from a four-year university. From the *Parents Education* graph, all bars highlighted as yellow are considered first gen, while all bars in blue are not (excluding Other/Unknown).

<br>

# Academic Characteristics

```{r, include=FALSE}
bns <- bns %>%
  mutate(school_split = fct_collapse(school,
                                     "CC" = c("Allan Hancock Community College",
                                              "Butte CC",
                                              "Clovis CC",
                                              "Mt. SAC CC",
                                              "Palo Verde CC"),
                                     "State" = c("Cal State LA",
                                                 "Cal State San Bernardino",
                                                 "CSU Bakersfield",
                                                 "CSU Dominguez Hills",
                                                 "Sacramento State",
                                                 "San Francisco State"),
                                     "UC" = c("UC Berkeley")))

#rename to be consistent with code book
bns <- bns %>% 
  mutate(school_rnm = recode(school, "Butte CC"="Butte College", 
                       "Mt. SAC CC" = "Mt. San Antonio College",
                       "Clovis CC" = "Clovis Community College",
                       "Palo Verde CC" = "Palo Verde College",
                   "Allan Hancock Community College" = "Allan Hancock College"))

```

The following set of questions were asked in regards to the students' academic status. This includes information such as what college they are currently enrolled in, their class standing, major, GPA, and enrolled units. They were also asked questions about programs they participate in as well as what type of student they are (transfer, international, etc.) to help identify students.


## School

**Which campus do you attend? `r print_n_reporting('school')`**

<div class = "row">

<div class = "col-md-7">

```{r}
#| fig-height: 6
#| fig-width: 8
plot_frq(bns$school_rnm, geom.colors = plot_color, ylim=c(0,450), hjust = -.2, vjust = "center",
         coord.flip = TRUE) + xlab('') + ylab("Number of Students")
```

Twelve campuses participated in this statewide survey. Of the twelve campuses, five were community colleges, six were California State Universities, and one was a University of California. The majority of students (`r unname(table(bns$school_split)[2])`) were currently enrolled in a state school. Fewer students (`r unname(table(bns$school_split)[1])`) reported enrollment in a community college, and a small minority (`r unname(table(bns$school_split)[3])`) reported enrollment in a UC. <br>

<br>

## Class Standing

**Estimate your class standing by units completed (do not include units from this semester/quarter): `r print_n_reporting('q15')`**

<div class = "row">

<div class = "col-md-7">

```{r}
#| fig-height: 4.5
#| fig-width: 5.5
tmp <- recode(bns$q15,
              "Freshman (0-29 semester units or 0-44 quarter units)" = "Freshman",
              "Sophomore (30-59 semester units or 45-89 quarter units)" = "Sophmore",
              "Junior (60-89 semester units or 90-134 quarter units)" = "Junior",
              "Senior (90+ semester units or 135+ quarter units)" = "Senior")
              
plot_frq(tmp, geom.colors = plot_color, wrap.labels = 8, ylim=c(0,750)) + xlab('Class Standing') + 
  ylab("Number of Students")
```

The statewide survey captures students of all class standing. Of the `r sum(table(bns$q15))` students, there were `r unname(table(tmp)[1])` freshmen, `r unname(table(tmp)[2])` sophomores, `r unname(table(tmp)[3])` juniors, and `r unname(table(tmp)[4])` seniors.

<br>

## Major

**Which of the following categories best describes your major? `r print_n_reporting('q16')`**

```{r}
#| fig-height: 6
#| fig-width: 8.5
bns <- bns %>% mutate(q16=recode(q16, 
                                 "Science, Technology, Engineering and Math (STEM) (Includes Medicine, Computer Science, Agriculture, Accounting, Statistics, Nursing, and Nutrition)" = "Science, Technology, Engineering and Math (STEM)"))

plot_frq(bns$q16, geom.colors = plot_color, ylim=c(0,875), wrap.labels = 16, sort.frq = 'desc') + xlab('') + ylab("Number of Students") + theme(axis.text.x = element_text(size = 10))

```

*NOTE: Science, Technology, Engineering and Math (STEM) was shortened for the figure above, but also includes the following majors: Medicine, Computer Science, Agriculture, Accounting, Statistics, Nursing, and Nutrition.*

The majority of students in the data are majoring in a `r names(sort(table(bns$q16), decreasing=TRUE)[1])` program (`r unname(sort(table(bns$q16), decreasing=TRUE)[1])` students). `r unname(sort(table(bns$q16), decreasing=TRUE)[2])` of the students are `r names(sort(table(bns$q16), decreasing=TRUE)[2])` majors.

<br>

## Overall GPA

**What is your approximate overall grade point average (GPA)? Remember that your GPA is on a scale from 0 to 4.0 `r print_n_reporting('q17')`**

```{r}
#| fig-height: 4
#| fig-width: 5.5
bns$gpa_fac <- cut(bns$q17, breaks = c(0, 1.5, 2.0, 2.5, 3.0, 3.5, 5.0), labels = c("1.5 or less", "1.51-2.0", "2.01-2.5", "2.51-3.0", "3.01-3.5", "3.51 or above"), ordered_result = TRUE, include.lowest = TRUE)

plot_frq(bns$gpa_fac, geom.colors = plot_color) + xlab('GPA') + ylab("Number of Students")
```

The average overall GPA for students is `r round(mean(bns$q17, na.rm = TRUE), 2)`, while the median is `r round(median(bns$q17, na.rm = TRUE), 2)`. The majority (`r percent(sum(prop.table(table(bns$gpa_fac))[5:6]), accuracy = .1)`) of students have a GPA greater than 3.0.

<br>

## Currently Enrolled Units

**How many units are you currently taking? `r print_n_reporting('q18')`**

```{r}
#| fig-height: 3.4
#| fig-width: 7
bns$q18 <- round(bns$q18, 0) # Rounded non-integers to integers
bns$units_binned <- cut(bns$q18, breaks = c(0, 6, 9, 12, 15, 18, 21, 1000), labels = c("6 or less", "7-9", "10-12", "13-15", "16-18", "19-21", "22+"), ordered_result = TRUE, include.lowest = TRUE)

plot_frq(bns$units_binned, geom.colors = plot_color) + xlab('Units') + ylab("Number of Students")
```

The average number of units a student is enrolled in is `r round(mean(bns$q18, na.rm = TRUE), 2)` units, while the median is `r median(bns$q18, na.rm = TRUE)` units. The majority of students are considered full-time (12 or more units), with `r names(sort(table(bns$q18), decreasing=TRUE)[1])` units being the highest occurring and `r names(sort(table(bns$q18), decreasing=TRUE)[2])` units being the second highest occurring.

<br>

# Student Identifiers

The following tables contain separate questions that were asked as a series of Yes/No responses to help further understand students. They were grouped together in this page by the start of the question.

## I identify as...

```{r}
questions <- c(paste0('q14_', 1:3), paste0('q14_', 5:12))
rnames <- c("A Transfer Student", "LGBTQA+", "A student with a disability", "A former foster youth/child",
            "An International student", "A non-California resident (not International)",
            "An ESL (English as a Second Language) student",
            "A Veteran of military service",
            "A student athlete",
            "A DREAM (i.e. undocumented) student",
            "A DACA (i.e. Deferred Action for Childhood Arrivals) student")
binary_table(questions, "Yes", rnames, ".")
```

<br>

## Disability Support Services

*Note: This name varies by campus and may not always be called the Disability Support Service Center.*

```{r}
#| fig-height: 5
#| fig-width: 7
questions <- paste0('q14_', 3:4)
rnames <- c('I identify as a student with a disability',
          'I receive services through my campus for a disability')
            
binary_table(questions, "Yes", rnames, "")
```
