---
title: "Food Security"
crossref:
  fig-title: 'Figure 6.'
  fig-labels: arabic
  title-delim: ":"
---

```{r, include=FALSE}
source("common_code.R")
```

The following questions were asked in regard to students' feelings towards food security due to a lack of or no money. Students were asked if they had to cut or skip meals (and how often), went entire days without eating, and other situations due to a lack of money.[^1] Students were also asked how they handled situations when they had run out of money for food, and if they had gotten any assistance.

[^1]: These questions were developed using a two-item version of the [USDA's screener for food insecurity status](https://www.ers.usda.gov/media/8279/ad2012.pdf)


## Food Eaten in Household

**For the following statements, please state whether they were often true, sometimes true, or never true for you/your household in the last 30 days.**

*Household is determined by the number of people you purchase and prepare food with. This may or may not include your roommates.*

```{r}
#| label: fig-q64-65
#| fig-cap: Food Eaten in Household
questions <- bns %>% select(q64, q65)
names(questions) <- c("I/We worried whether food would run out before I/we got money to buy more.",
                      "The food that I/we bought just didn’t last, and I/we didn’t have money to get more.")
plot_likert(questions,
            grid.range = c(1.2, 1.2),
            wrap.legend.labels = 20,
            expand.grid = FALSE,
            values = "sum.outside",
            catcount = 3,
            geom.colors = likert.pal.3,
            show.prc.sign = TRUE, reverse.scale = TRUE) + theme_bw() + theme(legend.position = "top")
```

<br>

## Food Insecurity

The following was determined using the two item food insecurity screening questions. If a student answered "sometimes true" or "often true" to either question, they were considered to be food insecure `r print_n_reporting("food_insecure_2item")`:

-   "I/We worried whether food would run out before I/we got money to buy more."
-   "The food that I/we bought just didn't last, and I/we didn't have money to get more."

::: row
::: col-md-7
```{r}
#| fig-height: 3.5
#| fig-width: 4.5
#| label: fig-two.item
#| fig-cap: Food Insecurity
bns <- bns %>% mutate(two.item = case_when(food_insecure_2item == 0 ~ "No",
                                           food_insecure_2item == 1 ~ "Yes",
                                           TRUE ~ as.character(food_insecure_2item)))
plot_frq(bns$two.item, geom.colors = plot_color, ylim=c(0,1250)) + xlab('Food Insecure') + ylab("Frequency")
```
:::

::: col-md-4
The majority of the students that took part in the survey are considered to be food insecure (`r unname(table(bns$two.item)['Yes'])` people). There are `r unname(table(bns$two.item)['No'])` students not considered to be food insecure, meaning they answered "never true" to the above question criteria.
:::
:::

<br>

## Ran Out of Money

**Additional food security question that are not part of the U.S. Adult Food Security Survey Module (10- item): In the last 30 days, if you ran out of money for food, what did you do? (choose all that apply):**


```{r}
#| label: fig-food_vars
#| fig-cap: Ran Out of Money
question <- "fs"
qlabel <- c("I did not run out of money for food",
            "My friends, family or roommates gave me money for groceries.",
            "My friends, family or roommates fed me or took me out to eat.",
            "I started skipping meals.",
            "I went to the food pantry or other free food organizations.",
            "I can get free food/meals from my work.",
            "I used my credit card (not debit card) to buy food.",
            "Other")

food_vars <- prep_binary_vars(question, qlabel) # %>% slice(-8) drop "other" since 0

# Plot table of percentages
food_vars %>% 
  select(Measure = xlab, "Yes (%)" = pct_lab) %>% 
  kable() %>% kable_styling(bootstrap_options = "striped")
```

